# Reproducible Research: Peer Assessment 1

```{r, echo=TRUE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, echo=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity_Data <- read.csv('activity.csv')
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r , echo=TRUE}
#activity_Data$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activity_Data$interval), format='%H:%M')
```

-----

## What is mean total number of steps taken per day?
```{r , echo=TRUE}
steps_By_Day <- tapply(activity_Data$steps, activity_Data$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r , echo=TRUE}
qplot(steps_By_Day, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r , echo=TRUE}
steps_By_Day_Mean <- mean(steps_By_Day)
steps_By_Day_Median <- median(steps_By_Day)
```
* Mean: `r steps_By_Day_Mean`
* Median:  `r steps_By_Day_Median`

-----

## What is the average daily activity pattern?
```{r , echo=TRUE}
average_Steps_Per_Time_Block <- aggregate(x=list(mean_Steps=activity_Data$steps), by=list(interval=activity_Data$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r , echo=TRUE}
ggplot(data=average_Steps_Per_Time_Block, aes(x=interval, y=mean_Steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r , echo=TRUE}
most_Steps <- which.max(average_Steps_Per_Time_Block$mean_Steps)
time_Most_Steps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", average_Steps_Per_Time_Block[most_Steps,'interval'])
```

* Most Steps at: `r time_Most_Steps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r , echo=TRUE}
num_Missing_Values <- length(which(is.na(activity_Data$steps)))
```

* Number of missing values: `r num_Missing_Values`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r , echo=TRUE}
activity_Data_Imputed <- activity_Data
activity_Data_Imputed$steps <- impute(activity_Data$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r , echo=TRUE}
steps_By_Day_Imputed <- tapply(activity_Data_Imputed$steps, activity_Data_Imputed$date, sum)
qplot(steps_By_Day_Imputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r , echo=TRUE}
steps_By_Day_Mean_Imputed <- mean(steps_By_Day_Imputed)
steps_By_Day_Median_Imputed <- median(steps_By_Day_Imputed)
```
* Mean (Imputed): `r steps_By_Day_Mean_Imputed`
* Median (Imputed):  `r steps_By_Day_Median_Imputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r , echo=TRUE}
activity_Data_Imputed$dateType <-  ifelse(as.POSIXlt(activity_Data_Imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r , echo=TRUE}
averaged_Activity_Data_Imputed <- aggregate(steps ~ interval + dateType, data=activity_Data_Imputed, mean)
ggplot(averaged_Activity_Data_Imputed, aes(interval, steps, color ="dateType"))        + geom_line() + 
         facet_wrap(.~dateType, nrow=2, ncol=1 ) +
         labs(x="5-minute interval",
         y="Avarage number of steps",
         title="Average daily steps by weektype")
```
